### 2018-07-02 papers/SiVL2018/objectnaming.tex
Adds first draft of structure; splits sections into files

\cite{graf2016animal}
\cite{Rosch1978}
\subsection{Intro to Problem}

\subsection{Definition Sub-, basic-, super-level Categories}
\cite{graf2016animal} investigate object naming with respect to reference level. They distinguish and manually annotate 3 levels: (i) sub-level (\textit{dalmatian}), (ii) basic-level (\textit{dog}), (iii) super-level (\textit{animal}).

\section{Requirements: What Do We Need? \cs{[Subsection of \ref{sec:object_naming}?]}}
For large-scale studies of object naming, we need to be able to automatically define the level of specificity of a name, given an ontology.