### 2018-07-07 papers/SiVL2018/objectnaming.tex
refines sec 2

%\cite{graf2016animal}
%\cite{Rosch1978}
%\subsection{Intro to Problem}
The act of naming an object amounts to that of picking out a nominal to be employed to refer to it (e.g., ``the \textit{dog}'', ``the white \textit{dog} to the left'') and can be seen as a subtask of generating a referring expression. The lexical choice involved in naming is a non-trivial one \cite{brown1958shall}. Indeed, since an object is simultaneously a member of multiple categories (e.g., a young beagle is at once a dog, a beagle, an animal, a puppy etc.), all the various names that lexicalize these constitute a valid alternative: in fact, their denotation includes the target object. The study of object naming targets those factors that lead to the choice of a particular name in communicative situations. These include properties of the object and candidate names, as well as contextual requirements.

Lexical alternatives in naming differ in their \textbf{level of specificity} (e.g., \textit{dog} is less specific than \textit{beagle})\cite{cruse1977pragmatics}: the categories and concepts they denote can indeed be organized in a hierarchical fashion according to class inclusion relations (e.g., a beagle is a dog, a dog is an animal etc.)  \cite{murphy2004big}. Such categories can be seen as giving rise to a taxonomy, or ontology, where locating an object at a certain category (e.g., beagle) allows to inductively access all its superordinate ones (e.g., dog, mammal, animal etc.).
It was observed that each type of object exhibits a preferred level of specificity which it is more naturally named with, called the \textbf{entry-level}. This typically corresponds to an intermediate level of specificity, i.e., \textbf{basic level} (e.g, \textit{bird}, \textit{car}) \cite{rosch1976basic}, as opposed to more generic (i.e., \textbf{super-level}; e.g., \textit{animal}, \textit{vehicle}) or specific categories (i.e., \textbf{sub-level}; e.g., \textit{sparrow}, \textit{convertible}). However, less prototypical members of basic-level categories tend to be instead identified with sub-level categories (e.g., a penguin is typically called a \textit{penguin} and not a \textit{bird}) \cite{jolicoeur1984pictures}.

Contextual factors also affect object naming. Scenarios where multiple objects are available induce a pressure for generating names which uniquely identify the target, thus excluding the competing alternatives (or \textit{distractors})\cite{olson1970language}. For example, in presence of more than one dog, the name \textit{dog} is ambiguous in terms of the specific one is intended to refer to. In these contexts, a sub-level category (e.g., \textit{rottweiler}, \textit{beagle}) is more informative, though possibly less default or more costly (e.g., infrequent, long etc.). It was shown that in use speakers tend to choose a name given a trade-off between its \textbf{cost}, on one side, and its contextual \textbf{ambiguity}, on the other; for example, they may opt for a less informative expression if less costly \cite{rohde2012communicating} \cite{graf2016animal}. Note that reasoning about the contextual informativeness of a name require categorization of all the potential referents, and not only the target object (e.g., to know that \textit{dog} is ambiguous in a context require having recognized that there are multiple dogs in the scene). Other contextual factors affecting lexical choice include the \textbf{perceptual salience} of the object, such as its size or prominence \cite{clark1983common}.

%\begin{itemize}
%\item Naming (lexical choice of nominal reference) is not trivial: many names for an object \cite{brown1958shall}. If seen as the act of picking out an applicable level: each object simultaneously belong to multiple categories, which reflect different conceptualizations and may be more or less appropriate across contexts,

% \cite{olson1970language}. Lexical alternatives: near-synonyms, hyponyms, hypernyms etc.\cite{edmonds2002near}

%\item Typically, variation across levels of specificity \cite{cruse1977pragmatics} Concepts are organized in a hierarchical fashion according to class inclusion relations, giving rise to a taxonomy \cite{murphy2004big}. Note that if accessing the most specific category in such taxonomy we can inherit knowledge about all the other categories it is also part of, by backtracking the path up to the root in the ontology (inductive access to lexical alternatives\begin{center}

%\item  Out-of-context preference for each object towards a certain specificity level, called \textbf{entry-level} category. Typically, this correspond to a intermediate level of specificity, called \textbf{basic level} \cite{rosch1976basic}, but for less typical members of the class may be more specific \cite{jolicoeur1984pictures}.

%\item Context: distractors, competing alternatives, pressure for unique reference. Overriding the out-of-context preference and may require the use of other levels of specificity e.g. \cite{murphy1989categorizing}. 
%In context, balancing a preference for low-cost, on one side, and informative, on the other, expressions: reduce cost while keeping interpretability (an ambiguous expression may be chosen if less costly (more frequent and short)) \cite{rohde2012communicating, graf2016animal}. In interaction: coherence and alignment. 
%\end{itemize}
%\subsection{Definition Sub-, basic-, super-level Categories}
%\cite{graf2016animal} investigate object naming with respect to reference level. They distinguish and manually annotate 3 levels: (i) sub-level (\textit{dalmatian}), (ii) basic-level (\textit{dog}), (iii) super-level (\textit{animal}).

\section{Requirements: What Do We Need? \cs{[Subsection of \ref{sec:object_naming}?]}}
For large-scale studies of object naming, we need to be able to automatically define the level of specificity of a name, given an ontology.