### 2018-07-09 papers/SiVL2018/proposal.tex

 vision people have focussed on learning very specific categories, and have not cared about learning actual concepts/names,

 language \& vision people have focussed on learning what speakers say in certain data sets, 
 we know that speakers will use object names at a medium level of specificity most of the time, 
 so it is difficult to learn the meaning of specific concepts from existing resources and to implement a model that can decide which level in the taxonomy is appropriate in context --- 
 we need a more systematic approach for collecting object naming data, 
 controlling the level of specificity (e.g. prompting speakers to come up with the most specific word they can think of)... this data might also give us insight into what basic-level and sub-level concepts are ..