### 2018-07-09 papers/SiVL2018/sivl_2018.tex
intermediate commit

% last updated in April 2002 by Antje Endemann
% Based on CVPR 07 and LNCS, with modifications by DAF, AZ and elle, 2008 and AA, 2010, and CC, 2011; TT, 2014; AAS, 2016

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{ruler}
\usepackage{booktabs}
\usepackage[usenames, dvipsnames]{color}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}

\newcommand{\sz}[1]{\textcolor{blue}{\emph{//sz: #1//}}}
\newcommand{\cs}[1]{\textcolor{PineGreen}{\emph{//cs: #1//}}}
\newcommand{\la}[1]{\textcolor{cyan}{\emph{//la: #1//}}}

\newcommand{\vgenome}{VisualGenome }
\newcommand{\referit}{ReferIt }
\newcommand{\refcoco}{RefCOCO }
\newcommand{\refcocop}{RefCOCO+ }
\newcommand{\flickr}{Flickr30k Entities }

\newcommand{\refexp}[1]{\textsl{#1}}
\newcommand{\cat}[1]{\textsc{#1}}

\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}

\begin{document}
% \renewcommand\thelinenumber{\color[rgb]{0.2,0.5,0.8}\normalfont\sffamily\scriptsize\arabic{linenumber}\color[rgb]{0,0,0}}
% \renewcommand\makeLineNumber {\hss\thelinenumber\ \hspace{6mm} \rlap{\hskip\textwidth\ \hspace{6.5mm}\thelinenumber}}
% \linenumbers
\pagestyle{headings}
\mainmatter
\def\ECCV18SubNumber{***}  % Insert your submission number here

\title{Object naming is context dependent - A case study on testing linguistic hypotheses on real-world image corpora} % Replace with your title

\titlerunning{ECCV-18 submission ID \ECCV18SubNumber}

\authorrunning{ECCV-18 submission ID \ECCV18SubNumber}

\author{Anonymous ECCV submission}
\institute{Paper ID \ECCV18SubNumber}

\maketitle

\begin{abstract}
The abstract should summarize the contents of the paper. LNCS guidelines
indicate it should be at least 70 and at most 150 words. It should be set in 9-point
font size and should be inset 1.0~cm from the right and left margins.
\dots
\keywords{We would like to encourage you to list your keywords within
the abstract section}
\end{abstract}

\section{Introduction}

Understanding and modeling the way humans converse about their environment using natural language has been a long-standing goal of research in various fields related to artificial intelligence and linguistics.
But, for a long time, computational research into language grounding, reference and situated interaction had to rely on ``block-worlds'' or very controlled, experimental environments \cite{anderson1991hcrc,fernangen:sigd07,krahmer:2012,takenobu2012rex,zarriess2016pentoref}, leading to models and corresponding findings at a rather restricted scale.
This situation has radically changed with recent developments in the language \& vision community. 
Pushed especially by developments in computer vision, there has been an explosion of interest in language \& vision tasks, ranging from image captioning \cite{fangetal:2015,devlin:imcaqui,chen2015mind,vinyals:show,Bernardietal:automatic}, referring expression resolution and generation \cite{Kazemzadeh2014,mao15,Yu2016,schlazar:acl16}, to multi-modal summarization or visual dialogue \cite{das2017visual,vries2017guesswhat}.

Despite this dramatic progress in the field, however, a lot of research in language \& vision still centers around systems for very specific tasks and data sets.

In this paper, we take a look at a straightforward task in language \& vision that so far has found surprisingly little attention in the community: object naming, i.e. determining the nominal that speakers will use to refer to an object in a visual scene. 
With state-of-the-art computer vision systems being able to classify images into thousands of different categories \cite{googlenet}, it might even seem like this task is already solved.
However, theoretically, the lexical choice involved in naming is a non-trivial one \cite{brown1958shall}.
Real-world objects are members of many categories and speakers can typically choose between different, more or less specific names when referring to a particular visual entity. 
For instance, the entity surrounded by the green box in Figure \ref{fig:example} is an instance of the categories \textit{female child, child, female, person, organism}, etc.\  and can be referred to with names such as, e.g.,\ \textit{girl, kid, cutie, daughter, person, human}.
Most research on computer vision and object recognition has somehow worked around the fact that objects can be instances of multiple, hierarchically embedded categories. 
State-of-the-art recognition models are mostly trained as classifiers on a flat of 1000 categories in ImageNet, reducing the ImageNet ontology to an inherently flat annotation scheme.

%But even though almost every task in language \& vision involves the prediction of object names (e.g. image captioning, referring expression generation, visual dialogue), hardly any research has explicitly looked at object naming, i.e.\ determining the actual word or concept that speakers would use to refer to an object (in a particular context), but see \cite{Ordonez:2016,zarriess-schlangen:2017}.
%Even research in pragmatics, that traditionally deals with reference and referring expression production, has mostly focused on attributes, rather than object names.
%Recently, however, \cite{graf2016animal} have shown that object naming preferences \textit{are} subject to pragmatic factors: in a typical reference game set-up with images of target objects, each  presented along with images of distractor objects, speakers have been found to flexibly adjust object names depending on the context. For instance, a \textit{dalmatian} would be called \textit{dalmatian} in the context of other dogs or simply \textit{dog} when none of the distractors is also a dog.
%This extends previous traditional work on concepts suggesting that the typicality of a referent determines its entry-level category (and consequently, its name) \cite{Rosch1978}.

The goal of this paper is to provide a systematic discussion of the object naming task and to show that existing, large-scale resources do not meet the requirements for a theoretically informed model of this very basic phenomenon in language \& vision.

\section{Naming Objects (in Context)}
\label{sec:object_naming}
\textbf{TODO@Laura}\\
\input{objectnaming}

\begin{figure}
	\centering
	\includegraphics[height=3.5cm]{fig/flickr_1000268201_boxes}
	\caption{INCLUDE MSCOCO examples}
	\label{fig:example}
\end{figure}

\section{Resources: What Do We Have?}
\label{sec:resources}
\textbf{TODO@Carina}\\
\input{resources}

\section{Proposal}
\label{sec:future_proposal}
\textbf{TODO@Sina (?)}\\
\input{proposal}

yes, will try ...

\clearpage
\bibliographystyle{splncs}
\bibliography{naming}
\end{document}