### 2019-02-02 papers/ACL2019/acl19_naming.tex
separated sections into files, some plans for analysis and reporting

\documentclass[11pt]{article}
%\usepackage[hyperref]{acl2017}
\usepackage[nohyperref]{acl2017}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}  %%% for including graphics
\usepackage{rotating}
\usepackage[dvipsnames]{xcolor}

%\aclfinalcopy % Uncomment this line for the final submission
\def\aclpaperid{489} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\newcommand{\sz}[1]{\textcolor{blue}{\emph{//sz: #1//}}}
\newcommand{\cs}[1]{\textcolor{Fuchsia}{\emph{//cs: #1//}}}
\newcommand{\gbt}[1]{\textcolor{green}{\emph{//gb: #1//}}}
\newcommand{\refexp}[1]{\textsl{#1}}
\newcommand{\word}[1]{\textsl{#1}}
\newcommand{\cat}[1]{\textsc{#1}}
\newcommand{\vgenome}{VisualGenome\xspace}
\newcommand{\ra}{$\rightarrow$}

%\title{What do you need to know about a word when you\\ want to use it for naming objects?}
%\title{How many times must an object be named (before we know what it is)?}

\title{Object naming on steroids}

\author{Sina Zarrie{\ss}  \and David Schlangen\\
  Dialogue Systems Group // CITEC // Faculty of Linguistics and Literary Studies \\
 Bielefeld University, Germany \\
  {\tt \{sina.zarriess,david.schlangen\}@uni-bielefeld.de} \\}

\date{}

\begin{document}
\maketitle

\begin{abstract}
\end{abstract}

\input{intro}
\input{relwork}
\input{data-collection}
\input{analysis}
%\input{modeling}
\input{conc}

\bibliography{naming}
\bibliographystyle{acl_natbib}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: