### 2019-02-02 papers/ACL2019/data-collection.tex

\section{Data Collection}
\label{sec:task}

%describe the YouNameIt task here

describe the task here

\subsection{Materials} describe sampling of images, category selection

Table with:

\begin{itemize}
\item rows: domains (if we go for long paper: then one row per collection node?)
\item columns:
  \begin{enumerate}
  \item \# collection nodes
  \item collection nodes (list)
  \item \# unique VG names
  \item example VG names
  \item \# unique objects
  \item \# unique images (? not sure if necessary; maybe only one of unique {objects, images})
  \end{enumerate}
\end{itemize}

\subsection{Procedure} describe the crowdsourcing set-up and the task

\subsection{Data} give an overview of the collected data

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "acl19_naming"
%%% End: