### 2019-05-19 papers/EMNLP2019/conclusion.tex

\subsection{Conclusion}
\label{sec:conclusion}

The object naming variation attested in the ManyNames dataset offers new challenges and perspectives, both for practical modeling approaches in Language \& Vision as well as for theoretical work in (Psycho-)linguistics.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: