### 2019-05-19 papers/EMNLP2019/analysis_table_wModelpreds.tex

\begin{figure*}
	\begin{minipage}[b]{0.5\linewidth}
		{\footnotesize
			\setlength{\tabcolsep}{1pt}
			\begin{tabular}{p{4cm}|p{4cm}|p{4cm}|p{4cm}}
				%\textbf{desk} &  \raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2320949_1048853_singleton_obj.png}} MN: keyboard  &
				%\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2343219_926143_supercat_unique.png}}  MN: desktop &
				%\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2354847_1742687_seed_ambiguous.png}} MN: computer \\
				%\textbf{bench} &  \raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2350360_1042111_supercat_unique.png}} MN: table  &
				%\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2389358_1261752_singleton_obj.png}}  MN: seat &
				%\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/1593011_2063521_singleton_obj.png}} MN: wood \\
				\multicolumn{4}{c}{\textbf{VG: sandwich}}\\
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2339876_3928476_supercat_unique.png}} sandwich (34) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2379889_1353176_supercat_unique.png}}  sandwich (15), basket (6), \underline{food} (5), burger (2),  \underline{hamburger} (2),  \underline{meal} (2) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2394266_465678_singleton_obj.png}} \underline{food} (10), sandwich (8), toast (5), french toast (4), dessert (2), breakfast (2) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2386509_681763_supercat_unique.png}} hotdog (14), \underline{food} (7), bun (4), sandwich (3),  \underline{bread} (2) [banana (1)]\\ 
				%
				 \textit{sandwich}, food, bread, sandwiches, plate, burger
				 & plate, \textit{food}, \textit{basket}, hotdog, wrapper, paper
				 & fish, chicken, meat, \textit{food}, \textit{sandwich}, bread
				 & \textit{hotdog}, \textit{bun}, \textit{banana}, \textit{bread}, table \\ 
				%
				%
				\multicolumn{4}{c}{\textbf{VG: bridge} }\\ 
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2341667_2006329_singleton_obj.png}} bridge (35)  &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/1592509_1610006_singleton_obj.png}} bridge (20),  \underline{building} (11)  &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2384683_1306430_singleton_obj.png}} street (16), road (15), bridge (3) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2412972_3494120_singleton_obj.png}} pier (6), railing (5), dock (5), bridge (5), fence (4), rail (3), boardwalk (3)\\
				\\ 
				%
				\textit{bridge}, arch, sky, structure, walkway
				& \textit{building}, \textit{bridge}, sky, structure, church
				& \textit{street}, \textit{road}, highway, intersection, pavement
				& \textit{railing}, water, \textit{pier}, ocean, \textit{fence}, deck, \textit{dock} \\ 
				%
				% 
				\multicolumn{4}{c}{\textbf{VG: bed}}\\ 
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2321254_3438076_singleton_obj.png}} bed (36)  &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2324306_3412337_singleton_obj.png}}  bed (16), bench (6), crib (5) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2342811_3485104_singleton_obj.png}}  bed (17), book (6), table (4), toy (3), bible (2), doll (2) & 
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/498222_3135415_singleton_obj.png}} bed (12), sleeping bag (9), blanket (7), bed sheet (5)\\ 
				%
				\textit{bed}, beds, sheets, frame, sheet
				& \textit{bed}, table, bed frame, sofa, couch
				& \textit{bed}, sheet, sheets, blanket, comforter, cover
				& \textit{bed}, comforter, \textit{blanket}, cover, sheet \\ 
				%
				% 
				\multicolumn{4}{c}{\textbf{VG: batter}}\\
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2372219_2683892_supercat_unique.png}} man (24), cap (5), \underline{person} (3), \underline{baseball player} (2) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2394377_464684_singleton_obj.png}} man (13), \underline{baseball player} (7), batter (5), \underline{player} (3), helmet (2) [player (1)]&
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2398907_2901496_singleton_obj.png}}  boy (7), helmet (5), \underline{baseball player} (4), \underline{player} (4), man (3), child (3), batter (3), dress (2), kid (2)&
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2337552_957263_singleton_obj.png}} pants (6), \underline{player} (5), shoe (4), bat (4), \underline{person} (4), legs (4), \underline{baseball player} (3), hitter (2)\\ 
				%
				\textit{man}, player, \textit{person}, shirt, guy
				& \textit{player}, \textit{batter}, \textit{man}, \textit{baseball player}, uniform
				& \textit{boy}, \textit{player}, \textit{batter}, \textit{kid}, girl, \textit{child}, \textit{baseball player}, person, uniform
				& \textit{player}, batter, \textit{person}, man, uniform, boy, \textit{baseball player}, dirt \\ 
				%
				% 
				\multicolumn{4}{c}{\textbf{VG: robe}}\\
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2373180_2333161_singleton_obj.png}} robe (27), \underline{bathrobe} (5) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/160_1058761_supercat_unique.png}} dress (30), uniform (2) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2334612_2838713_supercat_unique.png}}  robe (11), jacket (7), clothes (6), bag (3), dress (2), man (2) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2340041_2137546_supercat_ambiguous.png}} jacket (10), \underline{sweater} (9), coat (3), doll (3), toy (3), \underline{shirt} (2), bear (2), robe (2) [stuffed animal (1)]\\ 
				%
				woman, outfit, dress, girl, suit
				& woman, outfit, \textit{dress}, girl, suit
				& \textit{man}, shirt, person, \textit{robe}, coat, \textit{jacket}%, woman, outfit, clothes, clothing
				& teddy bear, \textit{bear}, dog, \textit{toy}, stuffed animal, blanket, toys
				%, stuffed animals \\ 
				%
				% 				
			\end{tabular}
		}
	\end{minipage}
	\caption{Examples for different instances of \vgenome synsets with low and high agreement in ManyNames. 
		First row: responses in MN, names that have a hierarchical relation to the \vgenome synset in WordNet are underlined.
	Second row: top name prediction by bottom up, names which match a MN respose are in italics.}
	\label{fig:ex-high-low-agreement}
\end{figure*}