### 2019-05-21 papers/EMNLP2019/conclusion.tex
modeling: finishes results+discussion; TODO: model+data formulation

\subsection{Conclusion}
\label{sec:conclusion}

The object naming variation attested in the ManyNames dataset offers new challenges and perspectives, both for practical modeling approaches in Language \& Vision as well as for theoretical work in (Psycho-)linguistics.

\cs{For future work, we want to use our data for fine-tuning ...; and model human object naming by training it on the full vocabulary of MN our names with a response count$>$n ... }

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: