### 2019-11-11 papers/LREC2020/analysis_table_wModelpreds.tex

\begin{figure*}[t]
	\begin{minipage}[b]{0.5\linewidth}
		{\scriptsize
			\setlength{\tabcolsep}{1pt}
			\begin{tabular}{p{2.6cm}|p{2.6cm}|p{2.6cm}|p{2.6cm}|p{2.6cm}|p{2.6cm}}
				%\textbf{desk} &  \raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2320949_1048853_singleton_obj.png}} MN: keyboard  &
				%\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2343219_926143_supercat_unique.png}}  MN: desktop &
				%\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2354847_1742687_seed_ambiguous.png}} MN: computer \\
				%\textbf{bench} &  \raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2350360_1042111_supercat_unique.png}} MN: table  &
				%\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2389358_1261752_singleton_obj.png}}  MN: seat &
				%\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/1593011_2063521_singleton_obj.png}} MN: wood \\
				\multicolumn{4}{c|}{\textbf{VG: sandwich}} & \multicolumn{2}{c}{\textbf{VG: robe}}\\
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2339876_3928476_supercat_unique.png}} sandwich (34) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2379889_1353176_supercat_unique.png}}  sandwich (15), basket (6), food (5), burger (2),  hamburger (2),  meal (2) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2394266_465678_singleton_obj.png}} food (10), sandwich (8), toast (5), french toast (4), dessert (2), breakfast (2) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2386509_681763_supercat_unique.png}} hotdog (14), food (7), bun (4), sandwich (3),  bread (2) [banana (1)] &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2373180_2333161_singleton_obj.png}} robe (27), bathrobe (5) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/160_1058761_supercat_unique.png}} dress (30), uniform (2)\\ \\
				%
				% \rowcolor{lightgray}
				%  \textit{sandwich}, food, bread, sandwiches, plate, burger
				%  & plate, \textit{food}, \textit{basket}, hotdog, wrapper, paper
				%  & fish, chicken, meat, \textit{food}, \textit{sandwich}, bread
				%  & \textit{hotdog}, \textit{bun}, \textit{banana}, \textit{bread}, table
				%  & woman, outfit, dress, girl, suit
				% & woman, outfit, \textit{dress}, girl, suit \\ 
				%
				%
				\multicolumn{4}{c|}{\textbf{VG: bridge} } & \multicolumn{2}{c}{\textbf{VG: robe}}\\ 
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2341667_2006329_singleton_obj.png}} bridge (35)  &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/1592509_1610006_singleton_obj.png}} bridge (20),  building (11)  &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2384683_1306430_singleton_obj.png}} street (16), road (15), bridge (3) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2412972_3494120_singleton_obj.png}} pier (6), railing (5), dock (5), bridge (5), fence (4), rail (3), boardwalk (3) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2334612_2838713_supercat_unique.png}}  robe (11), jacket (7), clothes (6), bag (3), dress (2), man (2) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2340041_2137546_supercat_ambiguous.png}} jacket (10), sweater (9), coat (3), doll (3), toy (3), shirt (2), bear (2), robe (2) [stuffed animal (1)]\\\\
				%
% 				\rowcolor{lightgray}
% 				\textit{bridge}, arch, sky, structure, walkway
% 				& \textit{building}, \textit{bridge}, sky, structure, church
% 				& \textit{street}, \textit{road}, highway, intersection, pavement
% 				& \textit{railing}, water, \textit{pier}, ocean, \textit{fence}, deck, \textit{dock} 
% 			         & \textit{man}, shirt, person, \textit{robe}, coat, \textit{jacket}%, woman, outfit, clothes, clothing
% 				& teddy bear, \textit{bear}, dog, \textit{toy}, \textit{stuffed animal}, blanket, toys
% \\ 
				%
				% 
				\multicolumn{4}{c|}{\textbf{VG: bed}} & \multicolumn{2}{c}{\textbf{VG: batter}}\\ 
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2321254_3438076_singleton_obj.png}} bed (36)  &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2324306_3412337_singleton_obj.png}}  bed (16), bench (6), crib (5) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2342811_3485104_singleton_obj.png}}  bed (17), book (6), table (4), toy (3), bible (2), doll (2) & 
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/498222_3135415_singleton_obj.png}} bed (12), sleeping bag (9), blanket (7), bed sheet (5) &
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2398907_2901496_singleton_obj.png}}  boy (7), helmet (5), baseball player (4), player (4), man (3), child (3), batter (3), dress (2), kid (2)&
				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2337552_957263_singleton_obj.png}} pants (6), player (5), shoe (4), bat (4), person (4), legs (4), baseball player (3), hitter (2)\\
				%
				% \rowcolor{lightgray}
				% \textit{bed}, beds, sheets, frame, sheet
				% & \textit{bed}, table, bed frame, sofa, couch
				% & \textit{bed}, sheet, sheets, blanket, comforter, cover
				% & \textit{bed}, comforter, \textit{blanket}, cover, sheet 
				% & \textit{boy}, \textit{player}, \textit{batter}, \textit{kid}, girl, \textit{child}, \textit{baseball player}, person, uniform
				% & \textit{player}, batter, \textit{person}, man, uniform, boy, \textit{baseball player}, dirt \\  
				%
				% 
%				\multicolumn{4}{c}{\textbf{VG: batter}}\\
%				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2372219_2683892_supercat_unique.png}} man (24), cap (5), person (3), baseball player (2) &
%				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2394377_464684_singleton_obj.png}} man (13), baseball player (7), batter (5), player (3), helmet (2) [player (1)]&
%				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2398907_2901496_singleton_obj.png}}  boy (7), helmet (5), baseball player (4), player (4), man (3), child (3), batter (3), dress (2), kid (2)&
%				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2337552_957263_singleton_obj.png}} pants (6), player (5), shoe (4), bat (4), person (4), legs (4), baseball player (3), hitter (2)\\ 
%				%
%				\rowcolor{lightgray}
%				\textit{man}, player, \textit{person}, shirt, guy
%				& \textit{player}, \textit{batter}, \textit{man}, \textit{baseball player}, uniform
%				& \textit{boy}, \textit{player}, \textit{batter}, \textit{kid}, girl, \textit{child}, \textit{baseball player}, person, uniform
%				& \textit{player}, batter, \textit{person}, man, uniform, boy, \textit{baseball player}, dirt \\ 
				%
				% 
%				\multicolumn{4}{c}{\textbf{VG: robe}}\\
%				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2373180_2333161_singleton_obj.png}} robe (27), bathrobe (5) &
%				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/160_1058761_supercat_unique.png}} dress (30), uniform (2) &
%				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2334612_2838713_supercat_unique.png}}  robe (11), jacket (7), clothes (6), bag (3), dress (2), man (2) &
%				\raisebox{-\totalheight}{\includegraphics[width=0.9\linewidth]{figures/2340041_2137546_supercat_ambiguous.png}} jacket (10), sweater (9), coat (3), doll (3), toy (3), shirt (2), bear (2), robe (2) [stuffed animal (1)]\\ 
%				%
%				\rowcolor{lightgray}
%				woman, outfit, dress, girl, suit
%				& woman, outfit, \textit{dress}, girl, suit
%				& \textit{man}, shirt, person, \textit{robe}, coat, \textit{jacket}%, woman, outfit, clothes, clothing
%				& teddy bear, \textit{bear}, dog, \textit{toy}, \textit{stuffed animal}, blanket, toys
%				%, stuffed animals \\ 
%				%
%				% 				
			\end{tabular}
		}
	\end{minipage}
	\caption{Examples for instances of \vgenome synsets with low and high agreement in ManyNames. 
		First row: responses in MN, names that have a hierarchical relation to the \vgenome synset in WordNet are underlined.
	Second row (gray): top name prediction by Bottom-up, names which match a MN response are in italics.}
	\label{fig:ex-high-low-agreement}
\end{figure*}