### 2020-03-30 papers/ACL2020/revision/introduction.tex

\begin{itemize}
\item We're interested in modeling human naming behavior [motivate]
\item names are sometimes used as labels for object classification in CV
\item usually: single-label evaluation scheme
  \begin{enumerate}
  \item if a single-label evaluation scheme is used, the target name should be the one that is most naturally used for the object. [Here we need a bit of background saying that it has been shown in Psycholing that people tend to agree in the name they give to objects.] To estimate this ``natural name'' (need better term), we need a reasonable sample.
  \item however, a single-label evaluation scheme assumes object names are unique (cf.\ evaluation; any other response is counted as incorrect). But alternative names are in fact possible.
  \end{enumerate}
\item (thanks to ManyNames,) we train models on the task of providing natural names; existing models, fine-tuned models 
\item and we analyze the consequences of evaluating them on a single-label scheme vs.\ one that allows for naming variants
\item (TBD: and we compare their errors/behavior to the human errors/behavior)
\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "acl2020_main"
%%% End: